### 2021-05-10 Manuscript/main.Rmd
Write manuscript

---
title: A demonstration of the \LaTeX class file for Statistics in Medicine with Rmarkdown
author:
- name: Boyi Guo*
  num: a
- name: Nengjun Yi
  num: a
address:
- num: a
  org: Department of Biostatistics, University of Alabama at Birmingham, Birmingham, USA
# - num: b
  # org: Department of Applied Things, University B, City B, Country B
corres: "*Corresponding author name, This is sample corresponding address. \\email{xxx@uab}"
presentaddress: This is sample for present address text this is sample for present address text
authormark: Guo and Yi
articletype: Research article
received: 2017-01-01
revised: 2017-02-01
accepted: 2017-03-01
abstract: "TODO: Change this accordingly; Lorem ipsum dolor sit amet, consectetur adipiscing elit. Aenean ut elit odio. Donec fermentum tellus neque, vitae fringilla orci pretium vitae. Fusce maximus finibus facilisis. Donec ut ullamcorper turpis. Donec ut porta ipsum. Nullam cursus mauris a sapien ornare pulvinar. Aenean malesuada molestie erat quis mattis. Praesent scelerisque posuere faucibus. Praesent nunc nulla, ullamcorper ut ullamcorper sed, molestie ut est. Donec consequat libero nisi, non semper velit vulputate et. Quisque eleifend tincidunt ligula, bibendum finibus massa cursus eget. Curabitur aliquet vehicula quam non pulvinar. Aliquam facilisis tortor nec purus finibus, sit amet elementum eros sodales. Ut porta porttitor vestibulum. Integer molestie, leo ut maximus aliquam, velit dui iaculis nibh, eget hendrerit purus risus sit amet dolor. Sed sed tincidunt ex. Curabitur imperdiet egestas tellus in iaculis. Maecenas ante neque, pretium vel nisl at, lobortis lacinia neque. In gravida elit vel volutpat imperdiet. Sed ut nulla arcu. Proin blandit interdum ex sit amet laoreet. Phasellus efficitur, sem hendrerit mattis dapibus, nunc tellus ornare nisi, nec eleifend enim nibh ac ipsum. Aenean tincidunt nisl sit amet facilisis faucibus. Donec odio erat, bibendum eu imperdiet sed, gravida luctus turpis."
keywords: Class file; \LaTeX; Statist. Med.; Rmarkdown;
bibliography: bibfile.bib
output: 
  rticles::sim_article:
    keep_tex: TRUE
---

\newcommand{\bs}[1]{\boldsymbol{#1}}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r child="01-Intro.Rmd"}
```

```{r child="02-Method.Rmd"}
```

```{r child="03-Simulation.Rmd"}
```

```{r child="04-Real_Data.Rmd"}
```

```{r child="05-Discussion.Rmd"}
```

```{r plot-ref, fig.cap = "Fancy Caption\\label{fig:plot}"}
x = rnorm(10)
y = rnorm(10)
plot(x, y)
```

...and can be referenced (Figure \ref{fig:plot}) by including the `\\label{}` tag in the `fig.cap` attribute of the R chunk: `fig.cap = "Fancy Caption\\label{fig:plot}"`. It is a quirky hack at the moment, see [here](https://github.com/yihui/knitr/issues/323).

Analogously, use Rmarkdown to produce tables as usual:

```{r, results = "asis"}
if (!require("xtable")) install.packages("xtable")
xt <- xtable(head(cars), caption = "A table", label = "tab:table")
print(xt, comment = FALSE)
```

## Cross-referencing

The use of the Rmarkdown equivalent of the \LaTeX cross-reference system
for figures, tables, equations, etc., is encouraged (using `[@<name>]`, equivalent of `\ref{<name>}` and `\label{<name>}`). That works well for citations in Rmarkdown, not so well for figures and tables. In that case, it is possible to revert to standard \LaTeX syntax.

# References
<div id="refs"></div>

# Appendix