### 2021-05-27 Simulation/tuning/Tuning Study Report.Rmd
Adding Simulation Study of Tuning Parameter

---
title: "Sim_tuning Report"
author: "Boyi Guo"
date: "5/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Computation environment

* Run on UAB Cheaha cluster. Find more description at [https://docs.uabgrid.uab.edu/wiki/Cheaha](https://docs.uabgrid.uab.edu/wiki/Cheaha)
* For the sake of time, the maximum number of iterations are set to 10.

# Restuls

* Poisson outcomes requires extremely more time to finish

* `bmlasso` doesn't not work well in gaussian outcomes. Similarly as we previously observed in netwrok ss studies

* Tuning parameter `s0` matters when the number of outcome grows exponentially.