### 2021-06-21 Real_Data/Emory_Card_Biobank/Report.Rmd
Add more content

---
title: "Emory Cardiovscular Biobank"
author: "Boyi Guo"
date: "5/31/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(targets)
library(readr)
library(here)
library(gtsummary)
library(tidyverse)
library(BhGLM)
library(BHAM)

train_dat <- targets::tar_read(RD_ECB_train_dat)
valid_dat <- targets::tar_read(RD_ECB_valid_dat)

gam_scn_res <- targets::tar_read(train_gam_screen)
```

#Background

Data was retrieved from <https://datadryad.org/stash/dataset/doi:10.5061/dryad.866t1g1mt> on May 31, 2021.

For study design, please visit the article <https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0237579>

## Data Introducetion

* Two data sets: model development set (N=`r NROW(train_dat)`), validation set (N=`r NROW(valid_dat)`)

* Variable Descriptions
GENEID: subject IDs
timetodeath3yr: time to death event or censoring at three years
death3yr: death event at three years
age: age in years
male: male gender, 1=male, 0=female
BlackRace: black race, 1=black, 0=non-black
batch: batch of metabolomics profiling
Strokehx: history of stroke, 1=yes, 0=no
CABGhx: prior CABG, 1=yes, 0=no
PVDhx: peripheral artery disease, 1=yes, 0=no
eGFR_max120_lt60: estimated glomerular filtration rate less than 60 ml/min/1.73m2, 1=yes, 0=no
curr_smoking: current smoking, 1=yes, 0=no
HFhx: heart failure history, 1=yes, 0=no
HTN: hypertension, 1=yes, 0=no
DM: diabetes, 1=yes, 0=no
mzXX_tXX: intensities of metabolic features

### Covariate Summary
```{r eval = F}
list(first_set = train_dat,
     second_set = valid_dat) %>% 
  map(~select(.x, # Data
              age, male, BlackRace,
           Strokehx:DM, 
           death3yr, timetodeath3yr) %>% 
        gtsummary::tbl_summary()
      ) %>% 
tbl_merge()
```

## Primary Data Analysis of Metabolites
```{r eval=F}
train_dat %>% select(starts_with("mz")) %>% 
  summarise_all(var)
```

## Screening
### Linear Effect Screening
```{r eval = F}
# feature_name <- 
linear_screening_res <- train_dat %>% select(starts_with("mz")) %>% 
  names() %>% 
map_dfr(#feature_name[1:2],
        .f = function(name, .dat){
  # name <- feature_name[1]
  y <- .dat %>% pull(death3yr)
  x <- .dat %>% pull({{name}})
  glm(y ~ x, family = binomial()) %>% 
    tidy() %>% 
    filter(term == "x") %>% 
    select(p.value) %>% 
    mutate(var = name,
           p.value)

},
.dat = train_dat)

lnr_scn_res <- data.frame(
linear_screening_res,
p.adj = p.adjust(linear_screening_res$p.value, "fdr")
) %>% arrange(p.adj) %>% filter(p.adj<0.2) #%>% 
  # slice_head(n = 100)
```

### GAM Screening
```{r eval = F}

library(mgcv)
library(broom)

# feature_name <- 
gam_screening_res <- train_dat %>% select(starts_with("mz")) %>% 
  names() %>% 
map_dfr(#feature_name[1:2],
        .f = function(name, .dat){
  # name <- feature_name[1]
  y <- .dat %>% pull(death3yr)
  x <- .dat %>% pull({{name}})
  mgcv::gam(y ~ s(x, bs = "cr", k = 5), family = binomial(),
            data = data.frame(x, y)) %>% 
    tidy() %>% 
    filter(term == "s(x)") %>% 
    select(p.value) %>% 
    mutate(var = name,
           p.value)

},
.dat = train_dat)

gam_scn_res <- data.frame(
gam_screening_res,
p.adj = p.adjust(gam_screening_res$p.value, "fdr")
) %>% arrange(p.adj) %>% #filter(p.adj<0.2)
  head(200)
```

# Predictive Model

Outcome: `death3yr` (bin)
Covariates: None
Predictors: `starts_with("mz")`

```{r data_split, eval = F}
library(tidymodels)

set.seed(1)

data_split <- rsample::initial_split(train_dat, prop = 3/4)

train_data <- training(data_split)
test_data <- testing(data_split)
```

## Baseline: Linear Lasso Model

```{r eval = F}
library(glmnet)

train_data <- train_dat %>% select(gam_scn_res %>% pull(var), death3yr)

lasso_mdl_cv <- cv.glmnet(x = train_data %>% select(starts_with("mz")) %>% data.matrix,
                    y = train_data %>% pull(death3yr), alpha = 1,
                    family = "binomial", type.measure = "auc")

lasso_mdl <- glmNet(x = train_data %>% select(starts_with("mz")) %>% data.matrix,
                    y = train_data %>% pull(death3yr), alpha = 1,
                    family = "binomial",lambda = lasso_mdl_cv$lambda.min)

lasso_cv_res <- cv.bh(lasso_mdl, ncv = 10)$measures

# class(lasso_mdl) <- c("glmNet", class(lasso_mdl))
# measure.bh(lasso_mdl)
# assess.glmnet(lasso_mdl$)

# measure.bh(lasso_mdl, new.x = test_data %>% select(starts_with("mz")) %>% as.matrix,
#            new.y = test_data %>% pull(death3yr))

# lasso_mdl <- glmnet(x = train_data %>% select(starts_with("mz")) %>% as.matrix,
#                     y = train_data %>% pull(death3yr), alpha = 1,
#                     family = "binomial",lambda = lasso_mdl_cv$lambda.min)
# 
# assess.glmnet(lasso_mdl, newx = train_data %>% select(starts_with("mz")) %>% as.matrix, newy = train_data %>% pull(death3yr))
# 
# assess.glmnet(lasso_mdl, newx = test_data %>% select(starts_with("mz")) %>% as.matrix, newy = test_data %>% pull(death3yr))

```

```{r}
library(glmnet)

lasso_mdl_cv <- cv.glmnet(x = train_dat %>% select(gam_scn_res %>% pull(var)) %>% data.matrix,
                    y = train_dat %>% pull(death3yr), alpha = 1,
                    family = "binomial", type.measure = "auc")

lasso_mdl <- glmNet(x = train_dat %>% select(gam_scn_res %>% pull(var)) %>% data.matrix,
                    y = train_dat %>% pull(death3yr), alpha = 1,
                    family = "binomial",lambda = lasso_mdl_cv$lambda.min)

lasso_cv_res <- cv.bh(lasso_mdl, ncv = 10)$measures

# class(lasso_mdl) <- c("glmNet", class(lasso_mdl))
# measure.bh(lasso_mdl)
# assess.glmnet(lasso_mdl$)

# measure.bh(lasso_mdl, new.x = test_data %>% select(gam_scn_res %>% pull(var)) %>% as.matrix,
#            new.y = test_data %>% pull(death3yr))

# lasso_mdl <- glmnet(x = train_data %>% select(starts_with("mz")) %>% as.matrix,
#                     y = train_data %>% pull(death3yr), alpha = 1,
#                     family = "binomial",lambda = lasso_mdl_cv$lambda.min)
# 
# assess.glmnet(lasso_mdl, newx = train_data %>% select(starts_with("mz")) %>% as.matrix, newy = train_data %>% pull(death3yr))
# 
# assess.glmnet(lasso_mdl, newx = test_data %>% select(starts_with("mz")) %>% as.matrix, newy = test_data %>% pull(death3yr))

```

```{r}
library(BHAM)

mz_names <- train_dat %>% select(gam_scn_res %>% pull(var)) %>% names  
sm_df <- data.frame(
  Var = mz_names,
 Func = "s",
 Args ="bs='cr', k=10"
)

sm_obj <- construct_smooth_data(sm_df, train_dat)
```
```{r}
dsn_mat <- sm_obj$data
```

```{r}
# The validation data doesn't overlap with training data
# test_sm_dat <- make_predict_dat(sm_obj$Smooth, dat = valid_dat)

# mdl_defaul <- bmlasso(dsn_mat, train_data$death3yr, family = "binomial", ss  = c(, 0.5))

bmlasso_mdl <- bmlasso(dsn_mat, train_dat$death3yr, family = "binomial",
                           group = make_group(names(dsn_mat)))
```

```{r}
spline_mdl <- bamlasso(dsn_mat, train_dat$death3yr, family = "binomial",
                           group = make_group(names(dsn_mat)))

# bmlasso_cv_res <- cv.bh(bmlasso_mdl, ncv = 10)$measures
# spline_cv_res <- cv.bh(spline_mdl, ncv = 10)$measures

# Training Data Performance
# cat("bmlasso model\n")
# measure.bh(mdl_defaul)
# cat("bmlasso_spline model\n")
# measure.bh(mdl_dflt)

```

```{r eval = F}
test_sm_dat <- BHAM::make_predict_dat(sm_obj$Smooth, dat=test_data)

# Testing Data Performance
cat("bmlasso model\n")
measure.bh(mdl_defaul, new.x = test_sm_dat, new.y=test_data$death3yr)
cat("bmlasso_spline model\n")
measure.bh(mdl_dflt, new.x = test_sm_dat, new.y=test_data$death3yr)
```

# Extremely Penalized Model
```{r}

mdl_pen_extreme <- bamlasso(dsn_mat, train_dat$death3yr, family = "binomial",
                           group = make_group(names(dsn_mat)),
                           ss = c(0.001, 0.5))

# spline_ext_cv_res <- cv.bh(mdl_pen_extreme, ncv = 10)

# Training Data Performance
# measure.bh(mdl_pen_extreme)

# Testing Data Performance
# measure.bh(mdl_pen_extreme, new.x = test_sm_dat, new.y=test_data$death3yr)
```

## Validation

The validation data set have mostly distinct feature from the training data set. Hence, the validation in the traditional sense does not exists.

```{r eval = F}

test_mz_names <- valid_dat  %>% select(starts_with("mz")) %>% names
measure.bh(mdl_defaul, new.x = test_sm_dat, new.y=valid_dat$death3yr)

measure.bh(mdl_dflt)
```