### 2021-10-01 Manuscript/00-main_ENAR.Rmd
Revise for ENAR submission

---
title: Spike-and-Slab Generalized Additive Models and Scalable Algorithms for High-Dimensional Data

# author: 
# - name: Author 1
  # affiliation: Department of YYY, University of XXX
  # email: abc@def

# - name: Author 2
  # affiliation: Department of ZZZ, University of WWW
  # email: djf@wef

keywords: Spike-and-Slab Priors; High-Dimensional Data; Generalized Additive Models; EM-IWLS; EM-Coordinate Decent; Scalablility

abstract: |
  Many proposed high-dimensional generalized additive models (GAMs) use sparse regularization, and tend to overly smooth important nonlinear functions. The excess shrinkage on coefficients can in turn produce inaccurate predictions. Moreover, most of these GAMs consider an “all-in-all-out” approach for function selection, rendering them difficult to answer if nonlinear effects are necessary. While some Bayesian models can address these shortcomings, using Markov chain Monte Carlo algorithms for model fitting creates a new challenge, scalability. Hence, we propose Bayesian hierarchical generalized additive models as a solution: we incorporate the smoothing penalty that can separate linear and nonlinear spaces of smoothing function. A novel spike-and-slab spline prior is used to select the components of smoothing functions. Two scalable and deterministic algorithms, EM-Coordinate Descent and EM-Iterative Weighted Least Squares, are developed for different utilities. Simulation studies and metabolomics data analyses demonstrate improved performance against state-of-the-art models, mgcv, COSSO and sparse Bayesian GAM. The software implementation is freely available via an R package BHAM.

bibliography: bibfile.bib
output: rticles::biometrics_article
# month: "`r format(Sys.Date(), '%b')`"
# year: "`r format(Sys.Date(), '%Y')`"
referee: true # Papers submitted to Biometrics should ALWAYS be prepared
              # using the referee option!!!! Turn off only to preview
              # two column-format
header-includes: 
  - \usepackage{tikz}
  - \usepackage{pgfplots}
  - \usepackage{amsfonts}
  # - \usepackage[modulo,pagewise,displaymath,mathlines]{lineno}
---

<!-- \usepackage{tikz} -->
<!-- \linenumbers -->
\pgfplotsset{compat=1.18}
\usetikzlibrary{shapes.geometric, arrows, positioning, calc, matrix, backgrounds, fit}
\newcommand{\bs}[1]{\boldsymbol{#1}}
\newcommand{\tp}{*}
\newcommand{\pr}{\text{Pr}}
\newcommand{\repa}{\text{repa}}
\newcommand{\simiid}{\overset{\text{iid}}{\sim}}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(knitr)
library(targets)
# library(here)
library(xtable)
library(flextable)

set_flextable_defaults(fonts_ignore=TRUE)

tar_load(intro_path)
tar_load(method_path)
tar_load(sim_path)
tar_load(real_data_path)
tar_load(disc_path)

tar_load(tab_binom)
tar_load(tab_gaus)

tar_load(WLM_train_dat)
# tar_load(tbl_one)
tar_load(ECB_bamlasso_cv)
tar_load(ECB_SBGAM_cv)
tar_load(WLM_bamlasso_cv)
tar_load(WLM_SBGAM_cv)
tar_load(tbl_real_time)
```

```{r child="01-Intro.Rmd"}
```
```{r child="02-Method.Rmd"}
```

```{r child="03-Simulation.Rmd"}
```

```{r child="04-Real_Data.Rmd"}
```

```{r child="05-Discussion.Rmd"}
```

\appendix
\section{}
\subsection{A.1 Derivation of $E({\tau^2}^{-1})$\label{ap:IWLS}}

\subsection{A.2 Weight Loss Maintenance Cohort Data Analysis\label{ap:WLM}}
We use the proposed models BHAM-SSL to analyze metabolomics data from a recently published study [@Bihlmeyer2021] on the association between metabolic biomarkers and weight loss, where the dataset is publically available [@Bihlmeyer2021_data]. In this analysis, we primarily focused on the analysis of one of the three studies included, weight loss maintenance (WLM) cohort [@Svetkey2008], due to the drastically different intervention effects. In the dataset, 765 metaboliltes in baseline plasma collected were profiled using liquid chromatography mass spectrometry. Quality control and natural log transformation were performed during metabolites data preparation. The outcome of interest was standardized percent change in insulin resistance, and hence modeled using a Gaussian model. After removing missing datapoints and addressing outliers in the data, there were `r nrow(WLM_train_dat)` remaining in the analysis. 5-Knot spline additive models for the gaussian outcome are constructed using two different models, the proposed BHAM-SSL and the SB-GAM. 10-Fold cross-validation(CV) are used to choose the optimal tuning parameters of each framework with respect to the default selection criterion implemented in the software. Out-of-bag samples were used for prediction performance evaluation, where deviance, mean squared error (MSE) defined as $\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2$, and mean abosolute error (MAE)defined as $\frac{1}{n}\sum\limits^{n}_{i=1}|y_i - \hat y_i|$ were calculated and tabulated in Table \ref{tab:WLM_res}.

```{r results = "asis"}
rbind(
  'BHAM-SSL' = WLM_bamlasso_cv %>% slice_min(mse) %>% select(-s0),
  'SB-GAM' = WLM_SBGAM_cv %>%  slice_min(mse) %>% select(-lambda)
) %>%
  rownames_to_column("Methods")  %>%
  xtable(
    caption = "Measures of optimal BHAM-SSL and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation",
    label = "tab:WLM_res"
  ) %>%
  print(comment = FALSE, include.rownames = FALSE, caption.placement = "top")

  # flextable() %>%
  # flextable::autofit() %>%
  # flextable::set_caption(caption = "Measures of optimal BHAM-SSL and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation\\label{tab:WLM_res}")
```

\newpage