### 2022-01-07 Manuscript/R&R/00-response_letter.Rmd
Add Review Comments

---
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\hrule
\vspace*{0.02cm}
\hrule
\vspace*{1cm}
\Large __Response to Reviewers' Comments__
\vspace*{1cm}
\hrule
\vspace*{0.5cm}
\begin{tabular}{llp{11.5cm}}
Manuscript No. & : & SIM-21-0956 \\
Title & : & Spike-and-Slab Generalized Additive Models and Scalable Algorithms for High-Dimensional Data\\
\end{tabular}
\vspace*{0.5cm}
\hrule
\vspace*{0.02cm}
\hrule

\vspace*{1cm}

First, we would like to thank you for handling this paper and thank the AE and both reviewers for their valuable comments. We have carefully addressed all the comments. Point-by-point responses are given below, where each comment is quoted in _italics_.

\bigskip

# Response to Associate Editor
_Comments to the authors:_

_Please follow carefully the comments, criticisms and suggestions of the three referees._

__Response__: We summarize the major changes of our manuscript as a response to address the comments from the referees:
\begin{itemize}
    \item xxx
    \item yyy
    \item zzz
\end{itemize}

\newpage
# Response to Reviewer \#1
```{r reviewer1, child="01-reviewer_1.Rmd"}
```

\newpage
# Response to Reviewer \#2
```{r reviewer2, child="02-reviewer_2.Rmd"}
```

\newpage
# Response to Reviewer \#4
```{r reviewer4, child="03-reviewer_4.Rmd"}
```

\newpage
# Reference