### 2022-01-07 Manuscript/R&R/03-reviewer_4.Rmd

We sincerely thank you for the valuable comments. We have carefully addressed all the comments. A point-by-point response is given below, where each of your comments is quoted in _italics_. __For a summary of major changes, please see our response to the Associate Editor.__

\bigskip

_Comments to the authors_

## _1. Summary_

_The authors introduce the Bayesian hierarchical generalized additive model (BHAM) for high-dimensional data analysis with generalized additive models (GAMs). The BHAM framework allows for variable selection and estimation in GAMs. However, in contrast to previous works on Bayesian GAMs, bilevel selection is possible under BHAM, and BHAM allows practitioners to more easily distinguish linear effects from nonlinear effects. Two deterministic algorithms, EM-Coordinate Descent (EM-CD) and EM-Iterative Weighted Least Squares (EM-IWLS), are introduced to fit BHAM. The EM-IWLS algorithm allows for inference, which is its main advantage over EM-CD. The algorithms appear to be faster than GAMs based on group regularization since fitting a model with a lasso penalty is typically faster than fitting one with a group lasso penalty._

_The paper is coherent and introduces a potentially promising methodology that addresses some of the shortcomings of existing Bayesian GAM approaches. However, in order to be suitable for publication in Statistics in Medicine, I have a few concerns that I think the authors should address. Namely, the authors need to do a better job illustrating the benefits of the BHAM approach over existing methods (such as the sparse Bayesian GAM method of [1, 2] or the spikeslabGAM of [3]). I detail my main concerns below._

## _2. Main Comments and Questions_
_The incorporation of the smoothing penalty matrix Sj in BHAM allows for separation of the linear space of a smoothing function from the nonlinear space. This is one of the touted benefits of BHAM - this separation allows for bilevel selection rather than "all-in-all-out" selection, so that the practitioner can distinguish linear effects from nonlinear ones. However, one concern is that the method might select the nonlinear effects but fail to select the linear component. The authors brie
y allude to this in the Conclusion (p. 13 of the manuscript). The authors say that they "\[include] the linear component in the model when non-linear component is selected."_

\em 
Could the authors elaborate on how they do this? I think this point should also be
stressed and made more explicit earlier in the manuscript, possibly in Section 2 where the EM-CD and EM-IWLS algorithms are introduced. One potential downside of
using the double exponential priors on individual basis coefficients, as opposed to the
group spike-and-slab prior in SBGAM, is the fact that only individual coefficients are
regularized. So it is conceivable that BHAM would end up including the nonlinear
component but not the linear one, which is not sensible. How can strong hierarchy
be enforced to ensure that the linear component is always included if the nonlinear
one is selected (but not necessarily the other way around)? This should be detailed in
the EM-CD and EM-IWLS algorithms, and the authors should explain how to enforce
strong hierarchy.

For example, in bilevel selection with the sparse group lasso [4], Simon et al. (2012)
state in their algorithm that they first check whether or not the group is selected,
and only if the group is nonzero do they then proceed with regularizing the individual
coordinates in the nonzero groups. Is a similar check being done for BHAM? This
needs to be explained further and made explicit. If necessary, the authors of the
present manuscript could include an additional subsection in Section 2 that explains
how to enforce strong hierarchy (i.e. linear effect is always selected if the corresponding
nonlinear effect for the jth covariate is selected).

2. Although the authors claim that BHAM is beneficial because it can distinguish linear
effects from nonlinear effects (therefore, we can tell if nonlinear effects are necessary),
this touted benefit is not well-illustrated in the Simulation Study (Section 3) or the
metabolomics data (Section 4). In particular, the simulation setting in Section 3.1
contains only one linear function. The authors should expand their simulation study
to include a greater number of linear functions (in addition to a few nonlinear ones)
to show how well BHAM is truly distinguishing the linear effects from the nonlinear
effects.

It would also be useful to report things like FDR, FNR, Matthews correlation coef-
ficient, etc. for the function selection to see how well the proposed BHAM model
is performing with respect to selection of the nonzero effects, especially compared to
other methods like SBGAM or spikeslabGAM. In addition, if the authors could re-
port a performance metric of how well BHAM is distinguishing nonlinear from linear
functions, e.g. possibly reporting the proportion of simulations where the linear func-
tion (resp. nonlinear) was correctly identified as linear (resp. nonlinear), that would
provide empirical support for BHAM over existing methods.

3. The authors claim that BHAM is computationally beneficial because it does not rely
on MCMC. The authors illustrate the computational benefit of the double exponential
prior over the group spike-and-slab prior of SBGAM [1, 2], but in my view, they
should also compare it to the method spikeslabGAM of [3]. There is an R package
spikeslabGAM to implement this method. This method uses a Gaussian spike-and-
slab prior on the basis coefficients and is implemented using MCMC. In order to assess
the benefit of using double exponential priors in BHAM vs. Gaussian spike-and-slab
priors, as well as illustrating the computational efficiency of the EM-CD and EM-
IWLS algorithms compared to MCMC, the authors should include comparisons to
spikeslabGAM.

4. The authors introduce the EM-IWLS method which can be used for inference of the
regression coefficients. Does this method work for p > n? Or is it only limited to the
case where p $\leq$ n? Even if this algorithm is limited to only the p $\leq$ n case, the benefit
of inference is not illustrated anywhere in the manuscript. Could the authors report
how to use the standard errors from the error variance-covariance matrix returned from
EM-IWLS to construct pointwise confidence intervals? There should be a section in
the manuscript for Uncertainty Quantification which explains how to use the results
from the EM-IWLS algorithm for the regression coefficients to construct pointwise
confidence intervals for the univariate functions themselves.

In addition, the coverage probability of the 95% pointwise intervals constructed by
BHAM in simulations should also be reported. Furthermore, the methods mgcv and
spikeslabGAM also return interval estimates, so the authors would do well to compare
the coverage probability of the confidence intervals constructed by BHAM vs. the
confidence/credible intervals returned by mgcv and spikeslabGAM.

5. The authors only provide one figure (Figure 1) which illustrates the hierarchical model
as a DAG. However, there are no other figures in the paper, e.g. plots of function
estimates from BHAM. The authors may consider adding a plot of the function esti-
mates vs. the ground truth in the simulation study (Section 3) for one of the nonlinear
functions and one of the linear functions (these could be, for instance, the left panel
and right panel of a figure). This will illustrate that BHAM can accurately recover
both nonlinear functions and linear functions.

The authors may also give one or two plots from the real data analysis in Section 4 of
some of the significant effects that they discovered. These plots should contain not just
the point estimates but also the the pointwise confidence intervals from the EM-IWLS
method. This will illustrate the benefit of EM-IWLS since it allows practitioners to
assess the uncertainty, in addition to an estimate of the function itself.

## _3. Minor Comments_
1. In Section 1, the authors say that nonparametric methods are an alternative to "black
box" methods like random forest. This is a bit misleading, since the "black box"
methods (random forest, deep learning, etc.) are als nonparametric methods. The
authors might consider rephrasing it (e.g. they could say that semiparametric methods
like GAMs allow for easier interpretation and selection of the significant effects than
the fully nonparametric "black box" methods).