### 2022-01-10 Manuscript/10-Supplementary.Rmd
Derivation for including linear & nonlinear selection hierarchy

---
title: "Supplementary File"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
\newcommand{\tp}{*}
\newcommand{\pr}{\text{Pr}}

## Supplementary 1: Marginal Distribution of $\gamma_j^\tp$
Given that $\gamma_{j}^\tp | \gamma_{j}, \theta_j \sim Bin(1, \gamma_{j}\theta_j)$ where $\gamma_{j} | \theta_j \sim Bin(1, \theta_j)$, we can derive the the marginal distribution of $\gamma_{j}^\tp$ with the following manipulation.
\begin{align*}
& \pr(\gamma_j^\tp = 1 | \theta_j)  = \pr(\gamma_j^\tp = 1, \gamma_j = 1 | \theta_j) + \pr(\gamma_j^\tp = 1 , \gamma_j = 0| \theta_j)\\
= & \pr(\gamma_j^\tp = 1, \gamma_j = 1 | \theta_j) + 0 \qquad \text{[hierarchical structure between }\gamma^\tp \text{ and }\gamma \text{.]}\\
= & \pr(\gamma_j^\tp = 1| \gamma_j = 1,  \theta_j)\pr(\gamma_j = 1| \theta_j)\\
= & \theta_j^2\\
& \pr(\gamma_j^\tp = 0 | \theta_j)  = \pr(\gamma_j^\tp = 0, \gamma_j = 1 | \theta_j) + \pr(\gamma_j^\tp = 0 , \gamma_j = 0| \theta_j)\\
= & \pr(\gamma_j^\tp = 0, \gamma_j = 1 | \theta_j) + \pr(\gamma_j^\tp = 0, \gamma_j = 0 | \theta_j)\\
= & \pr(\gamma_j^\tp = 0| \gamma_j = 1,  \theta_j)\pr(\gamma_j = 1| \theta_j) + \pr(\gamma_j^\tp = 0| \gamma_j = 0,  \theta_j)\pr(\gamma_j = 0| \theta_j)\\
= & (1-\theta_j)\theta_j + 1(1-\theta_j) = 1-\theta_j^2
\end{align*}