### 2022-02-11 Manuscript/R&R/01-reviewer_1.Rmd
Add R&R to the target pipeline; start to draft response.

We sincerely thank you for the valuable comments. We have carefully addressed all the comments. A point-by-point response is given below, where each of your comments is quoted in _italics_. __For a summary of major changes, please see our response to the Associate Editor.__

\bigskip

_Comments to the Authors_

_In this paper, the authors proposed a novel Bayesian hierarchical generalized additive model (BHAM) for high dimensional data analysis. Specifically, they incorporated smoothing penalties in the model via re-parameterization of smoothing function to avoid overly shrinking basis function coefficients. Incorporating the smoothing penalty allows the separation of the linear space of a smoothing function from the nonlinear space. Then they added a new two-part spike-and-slab spline prior on the smoothing functions for bi-level selection such that the linear and nonlinear spaces of smoothing functions can be selected separately. Two scalable optimization algorithms, EM-Coordinate Descent (EM-CD) algorithm and EM-Iterative weighted least square (EM-IWLS) algorithm, are developed and implemented._

_The proposed framework, BHAM, provides computational convenience using independent mixture double exponential distribution during model fitting by using optimization algorithms instead of intensive sampling algorithms and addresses the incapability of bi-level selection._

\bigskip

_Here are some comments._

_1. In the real data study, it is very important to establish the correlation between response variables and explanatory variables with appropriate models. The authors should add the comparison of linear models to illustrate that whether using non-linear was better. In addition, except expertise knowledge, are there metrics to judge whether a model is suitable for building model with non-linear part?_

__Response__: Thank you for the comment. We expanded our simulation study from two perspectives: 1) we added linear lasso models and spikeSlabGAM models, where linear lasso models serve as the reference to benchmark the performance of linear models; 2) we added new scenarios where the the functional form of each predictors are linear, to demonstrate the robustness of the proposed, regarding prediction and variable selection performances, when the non-linear assumption no longer holds. Please see the changes and conclusion in \emph{Section 3: SIMULATION STUDY}. To conclude, we observed [TODO: REPLACE HERE WITH CONCLUSION]. Hence, the proposed models perform better than linear models when functional are non-linear and as good as linear model when functions are linear. The expanded simulation study also demonstrates that the proposed model is flexible to model both linear and non-linear effects and broadly applicable in high-dimensional modeling without much constraints, which deems to be an additional strength.

\bigskip

_2. In Weight Loss Maintenance Cohort, R2 was very small by the two GAM methods.  I may think whether the data pretreatment and noise reduction process is appropriate._

__Response__:
\bigskip

_3. In the paper, several priors were mentioned. How to choose the prior in different situations? What is the advantage of the spike-and-slab double exponential prior?_

__Response__:
\bigskip

_4. How to detect the necessary before using the proposed model._

__Response__: