### 2022-02-15 Manuscript/R&R/00-response_letter.Rmd
Add simulaiton Settings

---
output: 
  pdf_document:
    keep_tex: FALSE
    citation_package: biblatex
bibliography: reply_ref.bib
header-includes: 
 - \usepackage{color}
---

\newcommand{\bg}[1]{\textcolor{red}{#1}}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(targets)
tar_load(RR_path)
```

\newcommand{\tp}{*}

\hrule
\vspace*{0.02cm}
\hrule
\vspace*{1cm}
\Large __Response to Reviewers' Comments__
\vspace*{1cm}
\hrule
\vspace*{0.5cm}
\begin{tabular}{llp{11.5cm}}
Manuscript No. & : & SIM-21-0956 \\
Title & : & Spike-and-Slab Generalized Additive Models and Scalable Algorithms for High-Dimensional Data\\
\end{tabular}
\vspace*{0.5cm}
\hrule
\vspace*{0.02cm}
\hrule

\vspace*{1cm}

First, we would like to thank you for handling this paper and thank the AE and both reviewers for their valuable comments. We have carefully addressed all the comments. Point-by-point responses are given below, where each comment is quoted in _italics_.

\bigskip

# Response to Associate Editor
_Comments to the authors:_

_Please follow carefully the comments, criticisms and suggestions of the three referees._

__Response__: We summarize the major changes of our manuscript as a response to address the comments from the referees:
\begin{itemize}
    \item According to Reviewer \#2's and Reviewer \#4's comments, we modified the prior distribution of the non-linear effects inclusion indicator such that the strong hierarchy for bi-level selection is enforced. 
    \item We moved the EM-IWLS fitting algorithm to the supporting information for three reasons. First of all, moving the EM-IWLS to the supporting information can help the audience focus on the model formulation and EM-Coordinate Descent algorithm, avoiding confusion and distraction. Secondly, many useful comments raised by the reviewers related to the EM-IWLS algorithm, e.g. constructing hypothesis testings and confidence band, requires a more dedicated discussion. We are preparing another manuscript discussing those. Lastly, relocating the EM-IWLS section would not hurt the integrity of the current manuscript that emphasize the scalability and bi-level selection.
    \item According to Reviewer \#1's and Reviewer \#4's comments, we added linear lasso model and spikeSlabGAM model among the models to be compared. 
    \item According to Reviewer \#1's and Reviewer \#4's comments, we added a new set of simulation scenarios where the functions of predictors are linear.  
    \item According to Reviewer \#4's comment, we summarized the variable selection performance for all models of comparison and bi-level selection performance for the models that is capable, i.e. the proposed model and spikeSlabGAM.
    \item According to Reviewer \#4's comment, we recorded the computation performance of spikeSlabGAM.
    \item All existing figure and tables were updated to reflect the methodology change and simulation design change. Figures were added according to Reviewer \#4's comment.  [TODO: think what figures are needed]
    \item We fixed the additional typos and grammar errors according to Review \#2. The revision of the manuscript is marked in \bg{red}. Deletions are not highlighted.
\end{itemize}

\newpage
# Response to Reviewer \#1
```{r reviewer1, child="01-reviewer_1.Rmd"}
```

\newpage
# Response to Reviewer \#2
```{r reviewer2, child="02-reviewer_2.Rmd"}
```

\newpage
# Response to Reviewer \#4
```{r reviewer4, child="03-reviewer_4.Rmd"}
```

\newpage
# Reference