### 2022-02-22 Manuscript/Tabs/sim_gaus_tab.tex

\begin{table}[ht]
\centering
\begin{tabular}{cccccccc}
  \hline
P & mgcv & LASSO & COSSO & Adaptive COSSO & BHAM & SB-GAM & spikeSlabGAM \\ 
  \hline
  4 & 0.90 (0.01) & 0.33 (0.01) & 0.71 (0.13) & 0.72 (0.11) & 0.90 (0.01) & 0.79 (0.04) & 0.80 (0.00) \\ 
   10 & 0.90 (0.01) & 0.33 (0.01) & 0.66 (0.21) & 0.77 (0.02) & 0.89 (0.01) & 0.79 (0.04) & 0.79 (0.00) \\ 
   50 & 0.86 (0.02) & 0.32 (0.01) & 0.46 (0.19) & 0.57 (0.18) & 0.80 (0.02) & 0.78 (0.05) & 0.78 (0.01) \\ 
  100 & NaN (NA) & 0.32 (0.01) & 0.41 (0.23) & 0.48 (0.25) & 0.79 (0.01) & 0.79 (0.05) & 0.77 (0.01) \\ 
  200 & NaN (NA) & 0.32 (0.01) & 0.39 (0.19) & 0.40 (0.17) & 0.79 (0.01) & 0.78 (0.04) & 0.75 (0.01) \\ 
   \hline
\end{tabular}
\caption{The average and standard deviation of the out-of-sample $R^2$ measure for
    Gaussian outcomes over 50 iterations. The models of comparison include the proposed Bayesian
    hierarchical additive model (BHAM) fitted with Iterative Weighted Least Square (BHAM-IWLS) and
    Coordinate Descent (BHAM-CD) algorithms, component selection and smoothing operator (COSSO), adaptive
    COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM). mgcv doesn't provide estimation
    whe number of parameters exceeds sample size i.e. p = 100, 200.} 
\label{tab:gaus}
\end{table}