### 2022-03-15 Manuscript/04-Real_Data.Rmd
Revise Real Data analysis

\section{Metabolomics Data Analysis}
\label{sec:real_data}

In this section, we apply the proposed models BHAM, fitted with the EM-CD algorithm, to analyze two real-world metabolomics datasets where the outcomes are binary and continuous respectively. We demonstrate the improved prediction performance compared to the other Bayesian hierarchical additive model, SB-GAM [@Bai2021], while being computationally efficient (see Table \ref{tab:time_real_data}). [TODO: add a sentenec how much faster.]
\begin{center}
Table \ref{tab:time_real_data} here
\end{center}

\subsection{Emory Cardiovascular Biobank}
\label{sec:ECB}
We use the proposed models BHAM to analyze a metabolic dataset from a recently published research [@Mehta2020] studying plasma metabolomic profile on the three-year all-cause mortality among patients undergoing cardiac catheterization. The dataset is publicly available via \textit{Dryad} [@Mehta2020_data]. It contains in total of 776 subjects from two cohorts. As there is a large number of non-overlapping features among the two cohorts, we use the cohort with larger sample size (N=454). There are initially 6796 features in the dataset, which is too large to be practically meaningful to analyze. Hence, we choose the the top 200 features \bg{with largest variance}. We use 5-knot spline additive models for binary outcome using two different models, the proposed BHAM and the SB-GAM. 10-Fold CV are used to choose the optimal tuning parameters of each framework with respect to the default selection criterion implemented in the software. Out-of-bag samples are used for prediction performance evaluation, where deviance, AUC, Brier score defined as $\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2$, and misclassication error defined as $\frac{1}{n}\sum\limits^{n}_{i=1}I(|y_i - \hat y_i|>0.5)$ are calculated. BHAM-CD obtains superior AUC, Brier score, and misclassification error in the out-of-bag samples compared to SB-GAM (see Table \ref{tab:ECB_res}).

\begin{center}
Table \ref{tab:ECB_res} here
\end{center}

\subsection{Weight Loss Maintenance Cohort}
We use the proposed models BHAM to analyze metabolomics data from a recently published study [@Bihlmeyer2021] on the association between metabolic biomarkers and weight loss, where the dataset is publicly available [@Bihlmeyer2021_data]. In this analysis, we primarily focus on the analysis of one of the three studies included, weight loss maintenance cohort [@Svetkey2008], due to the drastically different intervention effects. In the dataset, 765 metaboliltes in baseline plasma collected were profiled using liquid chromatography mass spectrometry. Quality control and natural log transformation \bg{were previously} performed and documented \bg{by the study publishing team [@Bihlmeyer2021]}. The outcome of interest are standardized percent change in insulin resistance, and hence modeled using a Gaussian model. After removing missing datapoints and addressing outliers in the data, there are $p$=`r nrow(WLM_train_dat)` features remaining in the analysis. 5-Knot spline additive models for the Gaussian outcome are constructed using two different models, the proposed BHAM and the SB-GAM. 10-Fold CV are used to choose the optimal tuning parameters of each framework with respect to the default selection criterion implemented in the software. Out-of-bag samples are used for prediction performance evaluation, where deviance, $R^2$, mean squared error (MSE) defined as $\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2$, and mean absolute error (MAE) defined as $\frac{1}{n}\sum\limits^{n}_{i=1}|y_i - \hat y_i|$ are calculated. BHAM obtains superior $R^2$, MSE, and MAE in the out-of-bag samples compared to SB-GAM (see Table \ref{tab:WLM_res}).
\begin{center}
Table \ref{tab:WLM_res} here
\end{center}