### 2022-03-21 Manuscript/03-Simulation.Rmd
Revise simulaiton results

\section{Simulation Study}
\label{sec:sim}

In this section, we compare the performance of the proposed model to \bg{six alternative models: linear LASSO models,} component selection and smoothing operator (COSSO) [@Zhang2006GAM], adaptive COSSO [@Storlie2011], generalized additive models with automatic smoothing (referred as \textit{mgcv} hereafter)[@Wood2011], \bg{spike-and-slab GAM \cite{Scheipl2012}}, and SB-GAM [@Bai2021]. \bg{We use linear LASSO model as the benchmark, examing the performance when linearity assumption doesn't hold}. COSSO is one of the earliest smoothing spline models that consider sparsity-smoothness penalty. Adaptive COSSO improved upon COSSO by using adaptive weight for penalties such that the penalty of each functional component are different for extra flexibility. \textit{mgcv} is one of the most popular models for nonlinear effect interpolation and prediction. Nevertheless, \bg{mgcv doesn't support analyses when the number of parameters exceeds the sample size.} \bg{Spike-and-slab GAM employs a spike-and-slab prior for GAM and uses a MCMC algorithm for model fitting.} SB-GAM is the first spike-and-slab LASSO GAM. We implement linear LASSO model with R package \texttt{glmnet 4.1-2}, COSSO and adaptive COSSO with R package \texttt{cosso 2.1-1}, generalized additive models with automatic smoothing with R package \texttt{mgcv 1.8-31}, spike-and-slab GAM with R package \texttt{spikeSlabGAM 1.1-15}, and SB-GAM with R package \texttt{sparseGAM 1.0}. COSSO models and SB-GAM do not provide flexibility to define smoothing functions, and hence use the default choices; mgcv, spikeSlabGAM and the proposed model allow customized smoothing functions and we choose the cubic regression spline. We control the dimensionality of each smoothing function, 10 bases, for all different choices of smoothing functions. We use 5-fold CV with the default selection criteria to select the final model for linear LASSO model, COSSO models, SB-GAM and the proposed model. 20 default candidates of tuning parameters ($s_0$ in BHAM, $\lambda_0$ in SB-GAM) are examined for SB-GAM and the proposed model which allow user-specification of tuning candidates. All computation was conducted on a high-performance 64-bit Linux platform with 48 cores of 2.70GHz eight-core Intel Xeon E5-2680 processors and 24G of RAM per core and R 3.6.2 [@R].

Other related methods for high-dimensional GAMs also exist, notably the methods of sparse additive models by Ravikumar et al. [@Ravikumar2009]. However, we exclude these methods from the current simulation study because of their demonstrated inferior predictive performance compared to \textit{mgcv} [@Scheipl2013].

\subsection{Monte Carlo Simulation Study}
We follow the data generating process described in Bai [@Bai2021]: we first generate $n=500$ training data points with $p=4, 10, 50, 100, 200$ predictors respectively, where the predictors $\bs X$ are simulated from a multivariate normal distribution $\text{MVN}_{n\times p}(0, I_{P})$. We then simulate the outcome $Y$ from two distributions, Gaussian and binomial with the identity link and logit link $g(x) = \log(\frac{x}{1-x})$ respectively. The mean of each outcome is derived via the following function
$$
\mathbb{E}(Y) = g^{-1}(5 \sin(2\pi x_1) - 4 \cos(2\pi x_2 -0.5) + 6(x_3-0.5) - 5(x_4^2 -0.3))
$$
for Gaussian and binomial outcomes. Gaussian outcomes requires specification of dispersion, where we set the dispersion parameter to be 1. In this data generating process, we have $x_1, x_2, x_3, x_4$ as the active predictors, while the rest predictors are inactive, i.e. $f_j(x_j) = 0$ for $j = 4, \dots, p$. Another set of independent sample of size $n_{test}=1000$, are created following the same data generating process, serving as the testing data. We generate 50 independent pairs of training and testing datasets to evaluate the prediction and variable selection performance of the chosen models, where training datasets are used to fit the models and testing datasets are used to calculate metrics of interest. \bg{In addition, we consider the data generating process where all functional forms of the predictors are linear while keeping the rest of simulation parameters the same. This additional set of linear simulations is designed to investigate the flexibility of the proposed model when nonlinear assumptions are not met.}

To evaluate the predictive performance of the models, the statistics, $R^2$ for Gaussian model and AUC for binomial model calculated based on the testing datasets, are averaged across 50 simulations. \bg{To evaluate the variable selection performance of the models, we record the set of variables each method selects and calculate the averaged positive predictive value (precision), true positive rate (recall), and Matthews correlation coefficient (MCC),
\begin{align*}
&\text{precision} = \frac{TP}{PP}\\
&\text{recall} = \frac{TP}{TP+FP}\\
&\text{MCC} = \frac{TP\times TN - FP \times FN}{\sqrt{(TP+FP)(TP+FN)(TN+FP)(TN+FN)}},
\end{align*}
where TP, TN, FP, FN, and PP are true positives, true negatives, false positives, false negatives, and predicted positive respectively. For the methods that don't automatically achieve variable selection, we set alpha level at 0.05 for \textit{mgcv} that relies on hypothesis testing, and a soft-threshold at 0.5 for spikeSlabGAM given the marginal inclusion probabilities. For the two methods, BHAM and spikeSlabGAM, that are capable of bi-level selection, we record the probability that the linear and nonlinear components of each predictors are selected in the models.}

\subsection{\bg{Simulation Results}}
\subsubsection{\bg{Prediction Performance}}
\bg{Among the set of simulations where the functional forms of the predictors are non-linear,} the predictive performances have a consistent pattern across the two distributions of outcomes. For simplicity, we use Gaussian simulations to exemplify the improvement of BHAM and defer to Tables \ref{tab:gaus} and \ref{tab:bin_auc} for detailed statistics. \bg{The proposed model, BHAM, predicts as good as, if not better than, other high dimensional additive models. Specifically, BHAM shows greate improvement over COSSO methods, resulting a median (interquartile range, IRT) 31\% (131\%) and 20\% (129\%) imporvement over COSSO and adaptive COSSO in $R^2$ statistics respectively. The improvement over spikeSlabGAM model is moderate, resulting in a median (IRT) 6\% (10\%) improvement in $R^2$. When comparing to SB-GAM, BHAM performs better (median (IRT) 13\% (8\%) improvement in R2) in lower dimensional cases ($p=4,10$), and equally good or slightly worse (median (IRT) 1\% (9\%) improvement in R2: ) in high-dimensional cases ($p=50,100, 200$). As previously hypothesized, the linear lasso model predicts less accurate compared to other flexible models across all scenario; mgcv performs extremely well in low-dimensional case ($p = 4, 10$), and deteriorates as the dimensionality increases until not applicable. To note, mgcv fits models but fails to converge within the default number of iterations when the sample size apporaches the number of coefficients to estimate ($p=50$), which leads to bad performance.} Even though SB-GAM has slight prediction advantage over the proposed model in high-dimensional situations, the BHAM has extreme computational advantage over SB-GAM, resulting median (IRT) 64\% (39\%) reduction in computation time (measured in seconds) for Gaussian simulations, without sacrificing much of the prediction accuracy (see Table \ref{tab:time_sim}).

\bg{We also examine the prediction performance when the functional form of predictors are linear, see Supporting Information Table S1, and S2. The proposed model, BHAM, has similar performance as the linear lasso model regardless of the distribution. This observation demonstrates that BHAM is a flexible model, and has good prediction performance regardless the underlying functional form of predictors. spikeSlabGAM have similar prediction performance to BHAM. Surprisingly, SB-GAM doesn't perform well in high-dimensional Gaussian outcome scenarios.}

\bg{\subsubsection{Variable Selection Performance}}
\bg{Among the set of simulations where the functional forms of the predictors are non-linear, the proposed model, BHAM, has a consistent performance across different dimension and distribution settings (See Table \ref{sec:gaus_var_select} for Gaussian outcomes, and Support Information Table S3 for binomial outcomes): being conservative. The symptoms of conservative variable selection are high precision and low recall, where high precision means that among all the selected variables, high percentage of them are true signals; low recall means that, the model selected small subset among all the active predictions. In other words, BHAM tends to select a smaller set of variables that are truly effective to the outcome. We want to note, the variable selection performance of BHAM is plummeted and not optimized when $p=200$. Upon further investigation, we discover it's because the generic sequence of $s_0$ used to tune the model doesn't contain the optimal value. Overall, among all the models examined, SB-GAM has the best performance, both high precision and high recall, and yield a high MCC. The performance of another Bayesian model, spikeSlabGAM deteriorates as the sparsity grows, particularly when ($p>50$), or for binomial outcomes. The variable selection performance for linear simulations match with prediction performance: BHAM performs great among the Gaussian scenarios, while the performance of SB-GAM deteriorates.}

\bg{Among the high-dimensional methods of comparison, there are two methods that are capable to achieve bi-level selection, the proposed BHAM and spikeSlabGAM. Among the linear simulations, both methods can accurately select the linear components and have a drastically lowered probability, close to 0, to include the nonlinear component, as anticipated. Specifically, spikeSlabGAM have smaller probability to include non-linear component in the model than BHAM. However, this advantage of spikeSlabGAM over BHAM is less obvious among the nonlinear simulations: spikeSlabGAM performs better than BHAM when selecting components of the function forms that include only linear or nonlinear component, e.g. function forms for $x_3$ and $x_4$. However, spikeSlabGAM inclines to ignore the variable that have more complex function forms, e.g. function forms for $x_1$ and $x_2$. In contrast, BHAM is more likely to include them in the model. This trade-off is determined by the assumption implicitly reflected via the prior hierarchy. We defer an in-depth discussion to the Section \ref{sec:concl}.}

<!-- #### Visualization -->
<!-- We plot the estimated smooth functions of the four active variables for the 100 simulation iterations, along with the 'true' smooth function. Across different settings of predictors, the patterns look similar. First of all, in all the settings, `mgcv` tends to provide more  extreme models, where in the graphs the smooth function have much higher bound than the true smooth function. Rarely the estimated smooth function from `mgcv` matches with the true smooth function. In comparison, the proposed method can interpolate the shape of the true smooth function. However, different priors and fitting algorithms have different advantages. For example, in Figure 1, we see that the SS normal distribution have problem with the tail of the smooth function, especially obvious in the estimations of the variable _x4_. However, it would have less sparse solutions, which is more likely to accurately interpolate the smooth function especially when the shape is complicated (see variable _x2_). On the contrary, the SS double exponential prior model encourages sparse solutions, which can sometime omit active predictors. -->