### 2022-03-21 Manuscript/Tabs/sim_binom_var_slct_tab.tex

\begin{table}[ht]
\centering
\begin{tabular}{cccccccc}
  \hline
P & Metric & LASSO & COSSO & Adaptive COSSO & BHAM & SB-GAM & spikeSlabGAM \\ 
  \hline
  4 & Precision & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 \\ 
   10 & Precision & 0.58 & 0.71 & 0.69 & 0.93 & 0.88 & 0.89 \\ 
   50 & Precision & 0.38 & 0.60 & 0.59 & 0.77 & 0.80 & 0.52 \\ 
  100 & Precision & 0.35 & 0.62 & 0.65 & 0.82 & 0.77 & 0.42 \\ 
  200 & Precision & 0.29 & 0.65 & 0.57 & 0.21 & 0.74 & 0.36 \\ 
    4 & Recall & 0.54 & 0.61 & 0.54 & 0.38 & 0.97 & 0.55 \\ 
   10 & Recall & 0.46 & 0.50 & 0.54 & 0.38 & 0.97 & 0.54 \\ 
   50 & Recall & 0.34 & 0.30 & 0.30 & 0.32 & 0.90 & 0.55 \\ 
  100 & Recall & 0.27 & 0.30 & 0.25 & 0.27 & 0.92 & 0.54 \\ 
  200 & Recall & 0.27 & 0.25 & 0.29 & 0.56 & 0.92 & 0.53 \\ 
   10 & MCC & 0.18 & 0.35 & 0.36 & 0.47 & 0.86 & 0.55 \\ 
   50 & MCC & 0.25 & 0.34 & 0.36 & 0.45 & 0.83 & 0.47 \\ 
  100 & MCC & 0.24 & 0.37 & 0.37 & 0.45 & 0.82 & 0.43 \\ 
  200 & MCC & 0.23 & 0.36 & 0.36 & 0.23 & 0.81 & 0.40 \\ 
   \hline
\end{tabular}
\caption{} 
\label{tab:sim_binom_var_select}
\end{table}