### 2022-03-21 Manuscript/Tabs/sim_lnr_binom_var_slct_tab.tex

\begin{table}[ht]
\centering
\begin{tabular}{cccccccc}
  \hline
P & Metric & LASSO & COSSO & Adaptive COSSO & BHAM & SB-GAM & spikeSlabGAM \\ 
  \hline
  4 & Precision & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 \\ 
   10 & Precision & 0.61 & 0.97 & 0.98 & 0.55 & 0.74 & 0.91 \\ 
   50 & Precision & 0.35 & 0.59 & 0.68 & 0.26 & 0.59 & 0.61 \\ 
  100 & Precision & 0.28 & 0.47 & 0.51 & 0.30 & 0.46 & 0.57 \\ 
  200 & Precision & 0.26 & 0.44 & 0.42 & 0.38 & 0.41 & 0.38 \\ 
    4 & Recall & 1.00 & 0.85 & 0.78 & 1.00 & 1.00 & 1.00 \\ 
   10 & Recall & 1.00 & 0.96 & 0.97 & 0.99 & 1.00 & 1.00 \\ 
   50 & Recall & 1.00 & 0.88 & 0.74 & 0.92 & 1.00 & 1.00 \\ 
  100 & Recall & 1.00 & 0.78 & 0.56 & 0.89 & 1.00 & 0.99 \\ 
  200 & Recall & 1.00 & 0.67 & 0.48 & 0.89 & 0.99 & 0.98 \\ 
   10 & MCC & 0.58 & 0.95 & 0.96 & 0.49 & 0.72 & 0.91 \\ 
   50 & MCC & 0.52 & 0.66 & 0.65 & 0.37 & 0.73 & 0.74 \\ 
  100 & MCC & 0.49 & 0.55 & 0.47 & 0.43 & 0.65 & 0.72 \\ 
  200 & MCC & 0.49 & 0.49 & 0.39 & 0.47 & 0.62 & 0.58 \\ 
   \hline
\end{tabular}
\caption{} 
\label{tab:sim_lnr_binom_var_select}
\end{table}