### 2022-04-02 Manuscript/R&R/00-response_letter.Rmd
correct typo

---
output: 
  pdf_document:
    keep_tex: FALSE
bibliography: reply_ref.bib
header-includes: 
 - \usepackage{color}
link-citations: true
---

\newcommand{\bg}[1]{\textcolor{red}{#1}}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(targets)
tar_load(RR_path)
```

\newcommand{\tp}{*}
\newcommand{\bs}{\boldsymbol}

\hrule
\vspace*{0.02cm}
\hrule
\vspace*{1cm}
\Large __Response to Reviewers' Comments__
\vspace*{1cm}
\hrule
\vspace*{0.5cm}
\begin{tabular}{llp{11.5cm}}
Manuscript No. & : & SIM-21-0956 \\
Title & : & Spike-and-Slab Generalized Additive Models and Scalable Algorithms for High-Dimensional Data\\
\end{tabular}
\vspace*{0.5cm}
\hrule
\vspace*{0.02cm}
\hrule

\vspace*{1cm}

First, we would like to thank you for handling this paper and thank the AE and both reviewers for their valuable comments. We have carefully addressed all the comments. Point-by-point responses are given below, where each comment is quoted in _italics_.

\bigskip

# Response to Associate Editor
_Comments to the authors:_

_Please follow carefully the comments, criticisms and suggestions of the three referees._

__Response__: We summarize the major changes in our manuscript as a response to address the comments from the referees:
\begin{itemize}
    \item According to Reviewer \#2's and Reviewer \#4's comments, we modified the prior distribution of the non-linear effects to reflect the effect hierarchy principle for bi-level selection. We updated the simulation and real data analysis results.
    \item We streamlined the structure of the manuscript, moving the previous EM-IWLS fitting algorithm subsection to the supporting information for three reasons. First of all, moving can help the audience focus on the model formulation and EM-Coordinate Descent algorithm, avoiding confusion and distraction. Secondly, many useful comments raised by the reviewers on the EM-IWLS algorithm, e.g. constructing hypothesis testings and confidence band, require a more dedicated discussion. We are preparing another manuscript discussing those. Lastly, with the added variable selection performance, the manuscript would be too lengthy if keeping the EM-IWLS in the main body. The readability of the manuscript would be damaged. Relocating the EM-IWLS section solves the dilemma without hurting the integrity of the current manuscript that emphasizes scalability.
    \item According to Reviewer \#1's and Reviewer \#4's comments, we added the linear LASSO model and spikeSlabGAM model among the models to be compared in the simulation study. In addition, we added a new set of simulation scenarios where the functions of predictors are linear.  
    \item According to Reviewer \#4's comment, we summarized the variable selection and bi-level selection performance in the simulation study section. The computation performance of spikeSlabGAM is recorded and addressed too.
    \item According to Reviewer \#4's comment, we added the plot showing the estimated functions of predictors for real-data analysis.
    \item We updated the feature screening procedure in our real-data analysis of the Emory Cardiovascular Biobank dataset. Our previous screening method, fitting univariate GAM, made the "double-dipping" mistake that would overestimate the prediction performance of any downstream prediction model. We updated the screening procedure with variance-based selection, and the prediction of both methods dropped. Nevertheless, the proposed model still outperforms the compared method. See change on P9.
    \item We fixed the additional typos and grammar errors according to Review \#2. The revision of the manuscript is marked in \bg{red} (roughly due to the technical difficulties associated with Latex). Deletions are not highlighted.
\end{itemize}

\newpage
# Response to Reviewer \#1
```{r reviewer1, child="01-reviewer_1.Rmd"}
```

\newpage
# Response to Reviewer \#2
```{r reviewer2, child="02-reviewer_2.Rmd"}
```

\newpage
# Response to Reviewer \#4
```{r reviewer4, child="03-reviewer_4.Rmd"}
```

\newpage
# Reference