### 2022-05-15 Manuscript/00-main.Rmd

---
title: Spike-and-Slab LASSO Generalized Additive Models and Scalable Algorithms for High-Dimensional Data Analysis
author:
- name: Boyi Guo*
  num: a
- name: Byron C. Jaeger
  num: b 
- name: AKM Fazlur Rahman
  num: a   
- name: D. Leann Long
  num: a  
- name: Nengjun Yi*
  num: a
address:
- num: a
  org: Department of Biostatistics, University of Alabama at Birmingham, Birmingham, USA
- num: b
  org: Department of Biostatistics and Data Science, Wake Forest School of Medicine, Winston-Salem, USA
corres: "Boyi Guo and Nengjun Yi, Department of Biostatistics, University of Alabama at Birmingham, Birmingham, USA. \\email{boyiguo1@uab.edu},  \\email{nyi@uab.edu}"
presentaddress: This is sample for present address text this is sample for present address text
authormark: Guo et al.
articletype: Research article
received: 2021-01-01
revised: 2021-02-01
accepted: 2021-03-01
abstract: "There are proposals that extend the classical generalized additive models (GAMs) to accommodate high-dimensional data ($p>>n$) using group sparse regularization. However, the sparse regularization may induce excess shrinkage when estimating smoothing functions, damaging predictive performance. Moreover, most of these GAMs consider an “all-in-all-out” approach for functional selection, rendering them difficult to answer if nonlinear effects are necessary. While some Bayesian models can address these shortcomings, using Markov chain Monte Carlo algorithms for model fitting creates a new challenge, scalability. Hence, we propose Bayesian hierarchical generalized additive models as a solution: we consider the smoothing penalty for proper shrinkage of curve interpolation via reparameterization. A novel two-part spike-and-slab LASSO prior for smoothing functions is developed to address the sparsity of signals while providing extra flexibility to select the linear or nonlinear components of smoothing functions. A scalable and deterministic algorithm, EM-Coordinate Descent, is implemented in an open-source R package BHAM. Simulation studies and metabolomics data analyses demonstrate improved predictive and computational performance against state-of-the-art models. Functional selection performance suggests trade-offs exist regarding the effect hierarchy assumption."
keywords: Spike-and-Slab Priors; High-Dimensional Data; Generalized Additive Models; EM-Coordinate Decsent; Scalablility; Predictive Modeling
# bibliography: bibfile.bib
output: 
  rticles::sim_article:
    keep_tex: TRUE
header-includes: 
  - \usepackage{tikz}
  - \usepackage{pgfplots}
  - \usepackage{pdflscape}
  - \usepackage{color}
  - \usepackage{soul}
---

<!-- \usepackage{tikz} -->
\pgfplotsset{compat=1.18}
\usetikzlibrary{shapes.geometric, arrows, positioning, calc, matrix, backgrounds, fit}
\newcommand{\bs}[1]{\boldsymbol{#1}}
\newcommand{\tp}{*}
\newcommand{\pr}{\text{Pr}}
\newcommand{\repa}{\text{repa}}
\newcommand{\simiid}{\overset{\text{iid}}{\sim}}
\newcommand{\bg}[1]{\textcolor{red}{#1}}
<!-- \newcommand{\st}[1]{} -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(knitr)
library(targets)
# library(here)
library(xtable)
library(flextable)

set_flextable_defaults(fonts_ignore=TRUE)

tar_load(manu_path)
tar_load(manu_tbs)
tar_load(manu_figs)

tar_load(WLM_train_dat)
tar_load(ECB_bamlasso_cv)
tar_load(ECB_SBGAM_cv)
tar_load(WLM_bamlasso_cv)
tar_load(WLM_SBGAM_cv)
tar_load(tbl_real_time)
tar_load(ECB_plot)
tar_load(WLM_plot)
```

```{r child="01-Intro.Rmd"}
```

```{r child="02-Method.Rmd"}
```

```{r child="03-Simulation.Rmd"}
```

```{r child="04-Real_Data.Rmd"}
```

```{r child="05-Discussion.Rmd"}
```

\clearpage

\bibliography{bibfile.bib}

\clearpage

\begin{figure}
\centering
\begin{tikzpicture} [
staticCompo/.style = {rectangle, minimum width=1cm, minimum height=1cm,text centered, draw=black, fill=blue!30},
outCome/.style={ellipse, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=blue!30},
mymatrix/.style={matrix of nodes, nodes=outCome, row sep=1em},
PriorBoarder/.style={rectangle, minimum width=5cm, minimum height=10cm, text centered, fill=lightgray!30},
background/.style={rectangle, fill=gray!10,inner sep=0.2cm, rounded corners=5mm}
]

\matrix (linearPrior) [matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
  \node (linearGamma) [outCome] { $\gamma_j \sim Bin(1, \theta_j) $ };\\
  \node (linearBeta) [outCome] { $\beta_j \sim DE(0,(1-\gamma_{j}) s_0 + \gamma_{j} s_1)$};\\
};
\matrix (penPrior) [right = 2cm of linearPrior, matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
  \node (penGamma) [outCome] { $\gamma_{j}^\tp \sim Bin(1, \bg{\gamma_{j}}\theta_j)$ };\\
  \node (penBeta) [outCome] { $\beta_{jk}^\tp \sim  DE(0,(1-\gamma^\tp_{j}) s_0 + \gamma^\tp_{j} s_1)$};\\
};

\node (s) [staticCompo]  at ($(linearBeta)!0.5!(penBeta)$)  {($s_0, s_1$)};
\node (Beta) [staticCompo, below = 1cm of s] {$\bs \beta = (\beta_1, \bs \beta^\tp_1, \dots,\beta_j, \bs \beta^\tp_j , \dots,\beta_p, \bs \beta^\tp_p) $};
\node (Theta)[outCome, above = 2cm of s] {$\theta_{j} \sim Beta(a, b)$};
\node (ab)[staticCompo, above = 0.5cm of Theta] {$(a, b)$};
\node (Y) [outCome, below = 1cm of Beta] {$y_i \sim Expo. Fam. (g^{-1}(\bs \beta^T \bs X_i), \phi)$};

\draw[->] (Theta) -- (linearGamma);
\draw[->] (Theta) -- (penGamma);
\draw[->] (linearGamma) -- (linearBeta) ;
\draw[->] (penGamma) -- (penBeta);
\draw[->, draw = red] (linearGamma) -- (penGamma);
\draw[->] (ab) -- (Theta);
\draw[->] (s) -- (linearBeta) ;
\draw[->] (s) -- (penBeta);
\draw[->] (linearBeta) -- (Beta);
\draw[->] (penBeta) -- (Beta);
\draw[->] (Beta) --  (Y);

\begin{pgfonlayer}{background}
  \node [background,
   fit=(linearGamma) (linearBeta),
   label=above:Linear Space:] {};
  \node [background,
    fit=(penGamma) (penBeta),
    label=above:Nonlinear Space:] {};
\end{pgfonlayer}

\end{tikzpicture}

\caption{Directed acyclic graph of the proposed Bayesian hierarchical additive model with parameter expansion. Elliposes are stochastic nodes, rectangles and are deterministic nodes. }
\label{fig:SSprior}
\end{figure}

\clearpage
\input{Tabs/sim_gaus_tab.tex}
\clearpage
\input{Tabs/sim_binom_tab.tex}
\clearpage
\input{Tabs/sim_time_tab.tex}
\clearpage
\input{Tabs/sim_gaus_var_slct_tab.tex}

\clearpage
<!-- \begin{landscape} -->

```{r sim_time, results = "asis", eval = F}
sim_tim_tab %>% 
  select(dist, p, `BHAM-IWLS`,`BHAM-CD`, cosso, acosso, mgcv, `SB-GAM`) %>% 
  rename(Distribution = dist,
         P = p,
         COSSO = cosso,
         `Adaptive COSSO` = acosso) %>% 
   xtable(
    caption = "The average and standard deviation of computation time in seconds, including cross-validation and final model fitting, over 50 iterations. The models of comparison include the proposed Bayesian hierarchical additive model (BHAM) fitted with Iterative Weighted Least Square (BHAM-IWLS) and Coordinate Descent (BHAM-CD) algorithms, component selection and smoothing operator (COSSO), adaptive COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM). mgcv doesn't provide estimation whe number of parameters exceeds sample size i.e. p = 100, 200.",
    label = "tab:time_sim",
    booktabs = TRUE
  ) %>% 
  print(comment = FALSE, include.rownames = FALSE,
       floating.environment = "sidewaystable",# caption.placement = "top"
       table.placement = "!h"
        )

  # flextable::flextable() %>% 
    # merge_v(j = 1) %>% 
  # width(j=1, width=1) %>% 
  # autofit() %>%
  # flextable::set_caption(caption = "The average and standard deviation of computation time in seconds, including cross-validation and final model fitting, over 50 iterations. The models of comparison include the proposed Bayesian hierarchical additive model (BHAM) fitted with Iterative Weighted Least Square (BHAM-IWLS) and Coordinate Descent (BHAM-CD) algorithms, component selection and smoothing operator (COSSO), adaptive COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM). mgcv doesn't provide estimation whe number of parameters exceeds sample size i.e. p = 100, 200.\\label{tab:time_sim}")
```
<!-- \end{landscape} -->
\clearpage

```{r }
tbl_real_time %>% 
  flextable::set_caption(caption = "Model fitting time in seconds for two metabolomics data analyses, from Emory Cardiovascular Biobank (ECB) and Weight Loss Maintenance Cohort (WLM). It tabulates the computation time for cross-validation step (CV) and optimal model fitting step (Final), and total computation time (Total) for the proposed model BHAM and the model of comparison SB-GAM.")
```
\label{tab:time_real_data}

\clearpage

```{r results = "asis"}
rbind(
  'BHAM' = ECB_bamlasso_cv %>% slice_max(auc)  %>% slice_min(s0) %>% select(-s0),
  'SB-GAM' = ECB_SBGAM_cv %>%  slice_max(auc) %>% select(-lambda)
) %>% 
  rownames_to_column("Methods") %>%
  select(-mae) %>% 
  rename(Deviance = deviance,
         AUC = auc,
         Brier = mse,
         Misclass = misclassification) %>% 
  xtable(
    caption = "Prediction performance of BHAM fitted with Coordinate Descent algorithm (BHAM) and SB-GAM models for Emory Cardiovascular Biobank by 10-fold cross-validation, including deviance, area under the curve (AUC), Brier score, and misclassification error (Misclass) where class labels are defined using threshold = 0.5.",
    label = "tab:ECB_res"
  ) %>% 
  print(comment = FALSE, include.rownames = FALSE#, caption.placement = "top"
        )
```

\clearpage

```{r results = "asis"}
rbind(
  'BHAM' = WLM_bamlasso_cv %>% slice_min(mse) %>% select(-s0),
  'SB-GAM' = WLM_SBGAM_cv %>%  slice_min(mse) %>% select(-lambda)
) %>% 
  rownames_to_column("Methods") %>%
  rename(
    Deviance = deviance,
    MSE = mse,
    MAE = mae,
    `$R^2$` = R2
  ) %>% 
    xtable(
    caption = "Prediction performance of BHAM fitted with Coordinate Descent algorithm (BHAM)  and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation, including deviance, $R^2$,  mean squared error (MSE), and mean absolute error (MAE).",
    label = "tab:WLM_res"
  ) %>%
  print(comment = FALSE, include.rownames = FALSE, sanitize.text.function=function(x){x}#, caption.placement = "top"
        )
  # flextable() %>% 
  # flextable::autofit() %>% 
  # flextable::set_caption(caption = "Measures of optimal BHAM and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation\\label{tab:WLM_res}")
```

\clearpage
\begin{figure}[h] 
\includegraphics{Figs/ECB_plot}
\caption{Plots of the functions for the 33 metablites selected by BHAM in the Emory Cardiovascular Biobank data analysis}
\label{fig:ECB_fig}
\end{figure}

<!-- \clearpage -->

<!-- \begin{figure}[h] -->
<!-- \includegraphics{Figs/WLM_plot} -->
<!-- \end{figure} -->