### 2022-05-15 Manuscript/01-Intro.Rmd

\section{Introduction}
\label{sec:intro}

Much modern biomedical research, e.g., sequencing data analysis and electronic health record data analysis, require special treatment of high-dimensionality, commonly known as $p >> n$ problem. There is extensive literature on high-dimensional linear models via penalized models or Bayesian hierarchical models, see Mallick and Yi [@Mallick2013] for review. These models are built upon a restrictive and unrealistic assumption, linearity. In classical statistical modeling, many strategies and models are proposed to relax the linearity assumption with various degrees of complexity. For example, variable categorization is a simple and common practice in epidemiology but suffers from power and interpretation issues. More complex models to address nonlinear effects include random forest and other so-called "black box" models [@Breiman2001]. These models are useful for statistical prediction but do not estimate parameters relevant to the data generation process that one can draw inferences from. In addition, how to generalize these \bg{"black box"} models to the high-dimensional setting remains unclear.

\bg{For their straightforward interpretation and flexibility, nonparametric regression models serve as great alternatives to the "black-box" models in prediction and variable selection}. Among those, generalized additive models (GAMs), proposed in the seminal work of Hastie and Tibshirani [@Hastie1987], grew to be one of the most popular modeling tools. In a GAM, the response variable, which is assumed to follow some exponential family distribution, can be modeled with the summation of smoothing functions. Nevertheless, the classical GAMs cannot fulfill the increasing analytic demands for high-dimensional data analysis.

There exist some proposals to generalize the classical GAM to accommodate high-dimensional applications. The regularized models, branching out from group regularized linear models, are used to fit GAMs by accounting for the structure introduced when expanding smoothing functions. Ravikumar et al. [@Ravikumar2009] extended the group LASSO [@Yuan2006] to additive models (AMs); Huang et al. [@Huang2010] further developed adaptive group LASSO for additive models; Wang et al. [@Wang2007] and Xue [@Xue2009] respectively applied group SCAD penalty [@Fan2001] to additive models. Bayesian hierarchical models are also used in the context of high-dimensional additive models, particularly within the spike-and-slab literature. Various group spike-and-slab priors [@Xu2015; @Yang2020] combining with computationally intensive Markov chain Monte Carlo (MCMC) algorithms are proposed, where the application on AMs is treated as a special case. Bai and co-authors[@Bai2020] were the first to apply group spike-and-slab LASSO prior to Gaussian AMs using a fast optimization algorithm and further generalized the framework to GAMs [@Bai2021]. Focus on addressing the sparsity, these methods can excessively penalize the bases of a smoothing function and produce inaccurate predictions, particularly when complex signals are assumed and large numbers of knots are used. [@Scheipl2013] In addition, these methods adopt an 'all-in-all-out' strategy, i.e. either including or excluding the variable completely, rendering no space for bi-level selection. Scheipl et al. [@Scheipl2012] proposed a spike-and-slab structure prior that addresses the bi-level selection. But the model fitting relies on computationally intensive MCMC algorithms and creates scalability concerns. Developing a fast, flexible and accurate generalized additive model framework would be of special interest.

\bg{We propose a novel Bayesian hierarchical generalized additive model (BHAM) for outcome prediction in the context of high-dimensional data analysis. Specifically, we incorporate smoothing penalties, derived from the smoothing spline literature \cite{Wood2017}, via reparameterization of smoothing functions to avoid excessive shrinkage on the bases. Smoothing penalties were also previously used in the spike-and-slab GAM \cite{Scheipl2012} and the sparsity-smoothness penalty \cite{Meier2009}. We then impose a new two-part spike-and-slab LASSO prior to address the signal sparsity. In addition, a scalable optimization-based algorithm, EM-Coordinate Descent (EM-CD) algorithm, is developed. While the primary focus of this model is to improve prediction, the proposed model also provides utility in functional selection. Notably, the two-part prior that follows the effect hierarchy principle motivates a bi-level selection, rendering one of three possibilities for each predictor: no effect, only linear effect, or linear and nonlinear effects. The proposed model is implemented in a publicly available R package \texttt{BHAM} via \url{https://github.com/boyiguo1/BHAM}.}

The proposed framework, BHAM, differs from previous spike-and-slab based GAMs, i.e., the spike-and-slab GAM [@Scheipl2012] and the sparse Bayesian GAM (SB-GAM) [@Bai2021], in three ways. Firstly, the proposed prior for smoothing functions is a spike-and-slab LASSO type prior using independent mixture double exponential distribution, compared to spike-and-slab GAM that uses normal-mixture-of-inverse gamma prior. Spike-and-slab LASSO priors provide computational convenience during model fitting by using optimization algorithms instead of intensive sampling algorithms. They make fitting high-dimensional models more feasible without sacrificing performance in prediction and variable selection. Secondly, SB-GAM uses a group spike-and-slab LASSO prior with an EM-CD algorithm to fit the model. While both methods use the combination of expectation maximization algorithm and coordinate descent algorithm, there are subtle differences in the implementation due to the difference in prior specification. The proposed model sets up independent priors among basis coefficients after the reparameterization step, which provides some advantage in computation. Lastly, the proposed model addresses the incapability of bi-level selection in SB-GAM.

In Section \ref{sec:BHAM}, we establish the Bayesian hierarchical generalized additive model, introduce the proposed spike-and-slab spline priors, and describe the fast-fitting EM-CD algorithm. In Section \ref{sec:sim}, we compare the proposed framework to state-of-the-art models via Monte Carlo simulation studies. Analyses of two metabolomics datasets are presented in Section \ref{sec:real_data}. Conclusion and discussions are given in Section \ref{sec:concl}.